---
title: "LVitovskyCaseStudy2Unit10"
author: "LVitovsky"
date: "July 22, 2016"
output: 
  html_document:
    keep_md: true
---
```{r set-options, echo=FALSE, cache=FALSE}
options(width = 80)

```

##Introduction

This paper seeks to asses the accuracy of three of the many forecasting techniques used by the research team of Amy Wheeler, Nina Weitkamp, Patrick Berlekamp, Johannes Brauer, Andreas Faatz, and Hans-Ulrich Holst (Osnabrück University, 2013) on the export data of three plants of the country "Chulwalar".

"Chulwalar"" is part of an island group "Urbano". The country is known for three types of plants that bloom in winter, the "Efak", the "Wuge", and the "Etel", which grows either blue or red flowers. This paper will focus on on one of these plants, the "Wuge".

The forecasting methods to be reviewed are:

  * Simple Exponential Smoothing
  
  * Holt's Linear Trend Method
  
  * Holt-Winters Seasonal Method (for both additive and multiplicative models)


##Understanding the Data
```{r ImportAndCleanRedo, include=FALSE}
source("~/testrepo/CaseStudy2/DataImportAndClean/ImportAndCleanRedo.R")
```
```{r EDA, include=FALSE}
source("~/testrepo/CaseStudy2/EDA/EDARedo.R")
```

The original data provided is a time series from 2008 to 2013, and includes not only export data for five plant groups (Efak, Wuge, and Etels, which were reported in total, blue, and red), but fourteen other parameters as well, such as births, national holidays, and inflation rate. 

After reproducing the data tidying, exploratory data analysis, and forecasting performed by the original researchers, here is summary of important points to know about the Wuge plant:
  
  * 1 - The raw data of the Wuge As-is data is not normally distributed, but performing a log transformation does help.
```{r histogram, echo=FALSE}
par(mfrow=c(1,2))
hist(WugeAsIs, col = "blue")
hist(log(WugeAsIs), col = "blue")

```

  * 2- A significance test of a linear relationship between the As-is Wuge Export Data and the Planned Wuge Export data produced a **p-value of <2e-16 with 70 degrees of freedom...**

  
```{r TTestOfCoefficients, echo=FALSE}

coeftest(WugeAsIs_tslm)

```

Because of these results, we can reject the null hypothesis that there is no linear relationship.  From the data analyzed, there is a slope.  This helps to support the strength of the correlation tests that followed between planned exports and as-is exports, as indicated in the next bullet point.

  * 3- Planned exports for the Wuge mirrored As-is exports very closely, with a **correlation of `r WugeCor`.**

  * 4- Correlations to each of the fourteen parameters are listed below:

```{r correlations, echo=FALSE}

#create a plot to show correlations side by side
par(mfrow=c(1,1), mar = c(12,6,4,2) + 0.1, mgp = c(4,2,0))
plot(Correlations, main="Correlations for Wuge Plant", xaxt="n", pch=19,col="blue", ylim=c(-1, 1), xlab = "", ylab="Correlation")
axis(1, labels = ParamNames, at=1:14, las=2)

```

Generally, the Wuge plant exports have the highest correlation to factors involving pricing and the overall health of the economy.  While one might think that births or National Holidays would correlate to Wuge exports, it is not clear if Chulwalar citizens gift these plants to each other, which would not be considered exports, anyway.  Also, the "Planned Spending for Independence Day Presents" could moreso be an indication of economic prosperity, not a direct link to the purchase of Wuge plants for those gifts specifically.  This could actually indicate that Chulwalar citizens are more *able* to purchase presents in general for their own family and friends. More research is needed regarding seasonality.

  * 5- After decomposing the Wuge export time series, one can see a gradual increasing trend, and that the variance of seasonality seems to increase over time. The remainder appears random, which is what we want in order for our three models to be as reliable as possible.

```{r STLPlots, echo=FALSE}
plot(WugeAsIs_stl, col="black", main="Decomposition of Wuge As-is")

```

##Accuracy of Forecasting

```{r forecasting, include=FALSE}
source("~/testrepo/CaseStudy2/ForecastsAndFindings/Forecasting.R")

```
###Simple Exponential Smoothing

Preliminarily, given the apparent upward trend and seasonality of Wuge exports, it is not likely that SES would prove to be a good method (https://www.otexts.org/fpp/7/1).  But, of course, let us be sure.  Below is plot of Wuge exports in black, with an SES overlay in green. The forecasted exports are in blue.

```{r SES, echo=FALSE}
par(mfrow=c(1,1))
Model_ses <- ses(WugeAsIs, h=12)
plot(Model_ses, plot.conf=FALSE, ylab="Wuge Exports Chulwalar", xlab="Year", main="", fcol="white", type="o")
lines(fitted(Model_ses), col="green", type="o")
lines(Model_ses$mean, col="blue", type="o")
legend("topleft",lty=1, col=c(1,"green"), c("data", expression(alpha == 0.497)),pch=1)
```

Quality measure generated were...

```{r SESQuality, echo=FALSE}
data.frame(QualityNames, SESMeasures)

```

Again, given the data, SES is not suspected to be the best fit, but it needs to be compared to the other models to get a relative sense of its worth.  We will keep moving to see if this method is better than the others...

###Holt's Linear Trend Method

The prospects of this method seem better than SES, given that Holt's can be used on data with a linear trend.  However, it might still not hold true given the seasonality of the Wuge.  Compare the side-by-side plots below.


```{r HoltsLinear, echo=FALSE}
par(mfrow=c(1,2))
plot(Model_holt_1, main="Holt's Additive Model")
plot(Model_holt_2, main="Holt's Exponential Model")

```

One can see that the forecast for the additive method (bold blue line) trend upward, with somewhat mirrored confidence and prediction intervals.  The multiplicative method has a more muted forecast for the next twelve months of Wuge exports, but more variance it seems with wider confidence and prediction intervals.

If we damp the methods so as to mute the effects of the data the further it is from the current day, our results are the following:

```{r HoltsLinearDamp, echo=FALSE}
par(mfrow=c(1,2))
plot(Model_holt_3, main="Holt's Additive w/ Damping")
plot(Model_holt_4, main="Holt's Multiplicative w/ Damping")

par(mfrow=c(1,1))
```

After dampening the trend to avoid too much optimism, the additive method showed a very slight gradual trend upward, and the exponential did the same. However, the exponential method includes much wider confidence and prediction intervals.  While we may have more conservative outlooks with the dampened method, we also are not as confident about forecasted results.

The quality measures generated were:

```{r HoltsQualityMeasures, echo=FALSE}
data.frame(QualityNames, HoltAdditive, HoltExponential, HoltAddDamp, HoltExpDamp)

```

From the numbers generated above, it appears that the Holt's Additive Model has the lowest AIC, AICc, and BIC, making it the best fit from these models.  Of course, there is still more to consider than just these numbers.

###Holt-Winters Seasonal Method

Given the seasonality of the Wuge exports, the expectation is that this method will prove to be more useful than the others. But of course, one must verify.  First we plot the forecasts...

```{r HoltWintersAdditiveAndExponential, echo=FALSE}
par(mfrow=c(1,2))
plot(Model_hw_1, main="Holt-Winters Additive")
plot(Model_hw_2, main="Holt-Winters Exponential")

```

The forecasts here are certainly much closer to the data visually, giving us an expected continued seasonality.  The quality measures for these two methods are below:


```{r HWAdditiveDF, echo=FALSE}
data.frame(QualityNames, HWAdditive, HWExponential)
```

Among these three quality measurements, the Holt-Winter's Exponential Method appears to be a better fit than the additive.  Given that the seasonality visually appears to increase over time, this makes sense.

Once again, we will damp the method to see if this improves the reliability of the forecasting provided.

```{r HWAdditiveAndExponentialDamp, echo=FALSE}
par(mfrow=c(1,2))
plot(Model_hw_3, main="Holt-Winters Add. w/ Damp")
plot(Model_hw_4, main="Holt-Winters Mult. w/ Damp")

```

From the plots, it appears that there is more variance in the forecast, confidence, and prediction intervals than in the non-damped methods. The plot below enables us to see all four models on top of each other, so see if the damped methods are more conservative (i.e., less optimisic). 


```{r HWPlots, echo=FALSE}
par(mfrow=c(1,1))
plot(HW1, main="Comparing Holt-Winters' models", plot.conf=FALSE, fcol="white", xlab="Year", lwd=2)
lines(fitted(HW2), col="purple", lwd=2)
lines(fitted(HW3), col="red", lwd=2)
lines(fitted(HW4), col="chartreuse", lwd=2)
lines(HW1$mean, col="black", lwd=2)
lines(HW2$mean, col="purple", lwd=2)
lines(HW3$mean, col="red", lwd=2)
lines(HW4$mean, col="chartreuse", lwd=2)
legend("topleft", lty=1, pch=1, col=c("black", "purple", "red", "chartreuse"), c("HW Additive", "HW Multiplicative", "HW Additive w/ Damp", "HW Multiplicative w/ Damp"))
```

When looking at the plot, the damped additive line (the red line) seems to have the highest end point, but with the least amount of variance.  All other lines seem to still have a good amount of variance comparatively.  Overall, all four lines seem to remain close to each other.

If we review the quality measures, and go ahead and place them next to the non-damped methods that we have already viewed, we have the following:

```{r HWQualityMeasuresWithDamp, echo=FALSE}
data.frame(QualityNames, HWAdditive, HWExponential, HWAdditiveDamp, HWExponentialDamp)

```


While dampening is usually preferred when forecasting data, the AIC, AICc, and BIC calculations tell us that they are not as reliable.  Perhaps the increasing seasonality is making it hard to get a good picture of when this will eventually decline.  Perhaps there has been an increased interest in the Wuge plant, and that will not alway be the case.  Of course, we have other metrics we can look at to decide if dampening should actually be used.



##So which one is "best"?

###AIC, AICc, and BIC

As previously states, when reviewing these quality measures, Holt-Winter's stood out as the clear frontrunner with the lowest values for AIC, AICc, and BIC.  More specifically, the Holt-Winter's multiplicative model was slightly lower than the Holt-Winter's additive model.

What is interesting is that while the Holt-Winter's model were undoubtedly better than the Simple Exponential Smoothing or Holt Linear Method, the difference between the additive and multiplicative is not as large.  Also, dampening the Holt-Winter model did not actually deliver better measures here.

###Looking at Error

Perhape reviewing the errors will give more clarity. Seven types of error are outlined below:

```{r Error, echo=FALSE}
data.frame(Methods, ME, RMSE, MAE, MPE, MAPE, MASE, ACF1)
```
Generally, the four Holt-Winters models have the smallest errors.  Their Mean Absolute Errors (MAEs) and Mean Absolute Percentage Errors (MAPEs) are about half of the rest of the methods.  The Mean Errors and Root Mean Squared Errors are lower for Holt-Winters as well, but given that they can magnify larger errors (Which there may be for this data given the apparent variance), more trust is placed on the absolute errors.

```{r ImportantErrors, echo=FALSE}
HWErrors

```


If choosing between a dampening method and not, it appears that the dampened method has more error tied to it.  Perhaps the increasing variance of the seasonality is affecting the ability of the model to provide confidence that those forecasts will materialize.

Between the non-damped Holt-Winters models, the additive model actually had less error associated with it in most cases.  Except for the Root Mean Squared Error, the exponential model had larger errors associated with it (albeit small), despite looking like the better fit when reviewing the quality measures (AIC, etc). So this counters our findings from analyzing the quality measures above.  Perhaps the additive model and exponential models are extremely close.

###Residuals

Investigating the residuals, a plot helps us compare the two models side by side:

```{r HWStates, echo=FALSE}
plot(states, xlab="Year")

```

The additive and multiplicative models are extremely close again.  Both have a steady increasing trend, as evidenced by the simple upward line in the "Level" plots.  The slopes are very similar, but still different enough to notice.  It appears that the additive model has a slight "v" shape, almost like a check mark, while the multiplicative model has more of an "L" shape. So the slope seems to increase more over time in the additive model, while it changes less over time in the multiplicative model.

The seasonality is overwhelming in this data.  Perhaps the strength of the seasonality is keeping us from getting a clear answer on which version, additive or multiplicative, of Hunt-Winters' Seasonal model is a better fit.

If we calculate the means and standard deviations of the residuals, we receive the following:
  
  * Means (added damped models just to be sure): 
    
    HW Additive Model: `r mean(Model_hw_1$residuals)`
    
    HW Multiplicative Model: `r mean(Model_hw_2$residuals)`
    
    HW Additive w/ Damp: `r mean(Model_hw_3$residuals)`
    
    HW Multiplicative w/ Damp: `r mean(Model_hw_4$residuals)`
    
    
  * Standard Deviations:
  
    HW Additive Model: `r sd(Model_hw_1$residuals)`
    
    HW Multiplicative Model: `r sd(Model_hw_2$residuals)`
    
    HW Additive w/ Damp : `r sd(Model_hw_3$residuals)`
    
    HW Multiplicative w/ Damp: `r sd(Model_hw_4$residuals)`
    
    
Here we see that dampening does not seem to be improving our confidence in the forecasts.  The standard deviations of the dampened models are actually larger, with a larger mean.  There is still conflicting information about additive versus multiplicative, however.  This warrants looking at plots to be sure the visual representation matches the statistics.
    

```{r ResidualPlots, echo=FALSE}
par(mfrow=c(1,2))
plot(Model_hw_1$residuals, main="Additive Model Residuals", xlab="Year", col="blue")
plot(Model_hw_2$residuals, main="Multiplicative Model Residuals", xlab="Year", col="blue")


```

From the side-by-side plots, it appears that the variance in the additive model continues to be a factor, while the multiplicative models is slightly more contained.  Both still have obvious variance, but starting in 2012, it appears that the downward movement of the residuals is not as deep as the additive's.  



##Conclusion

One can easily see that the Holt-Winters' Seasonal Model of forecasting is the best fit of the three tested forecasting techniques.  When choosing to use a dampening or not, given that this does not improve the reliabilty of the data, it is better to keep things simple and not dampen.

However, what is not so clear is if an additive or a multiplicative model should be used.  Taking all of the information presented here into account, it is recommended to use the **additive model**.   It is very close to the mutliplicative results, however, it had slightly **less error** tied to it.  Also, even though the slope of the residuals seemed to be more volatile than the multiplicative, perhaps this is just the true picture of the data given its smaller error measures. To have noticeably more change than another model, but to have less error, indicated to me that the model is a better fit.  It essentially had a "tough" job to do.

Going forward, forecasting for the Wuge plant will be challenging, as its seasonlity appears that it will continue to grow exponenentially.  From a business perspective, this could make the reliability of income from the Wuge less reliable.  Should something cause a high season to be de-railed, it could mean either 1) not being able to meet demand, or 2) having too much Wuges on their hands!

##Appendix

Code used to generate this document is below:

```{r Code, echo=TRUE}

#load correct libraries and the data
#download data, which is in two .csv files, AsIs data and Planned data
#set working directory to proper folder to read in .csv files

#load packages and data
setwd("~/testrepo/CaseStudy2/ChulwalarCase")
library(fpp) # for time series forecasting and analysis
library(forecast) # for some other forecasting models
ImportedAsIsData <- read.csv(file="ImportedAsIsDataChulwalar.csv", 
                             header = F, sep=";", fill = T)
ImportedIndicators <- read.csv(file="ImportedIndicatorsChulwalar.csv", 
                               header = F, sep=";", fill = T)
ImportedPlanData <- read.csv(file="ImportedPlanDataChulwalar.csv", 
                             header = F, sep=";", fill = T)

#check out the data
str(ImportedAsIsData)
str(ImportedPlanData)
str(ImportedIndicators)

#convert data into vectors
TotalAsIsVector <- c(ImportedAsIsData [2:13,2],
                     ImportedAsIsData [2:13,3],
                     ImportedAsIsData [2:13,4],
                     ImportedAsIsData [2:13,5],
                     ImportedAsIsData [2:13,6],
                     ImportedAsIsData [2:13,7])
WugeAsIsVector <- c(ImportedAsIsData [30:41,2],
                    ImportedAsIsData [30:41,3],
                    ImportedAsIsData [30:41,4],
                    ImportedAsIsData [30:41,5],
                    ImportedAsIsData [30:41,6],
                    ImportedAsIsData [30:41,7])
YearAsIsVector <- c(ImportedAsIsData [86,2],
                    ImportedAsIsData [86,3],
                    ImportedAsIsData [86,4],
                    ImportedAsIsData [86,5],
                    ImportedAsIsData [86,6],
                    ImportedAsIsData [86,7])
TotalAsIsVector_2014 <- c(ImportedAsIsData[2:13,8])
PlanVector <- c(ImportedPlanData[2:13,2],
                ImportedPlanData[2:13,3],
                ImportedPlanData[2:13,4],
                ImportedPlanData[2:13,5],
                ImportedPlanData[2:13,6],
                ImportedPlanData[2:13,7])
WugePlanVector <- c(ImportedPlanData[30:41,2],
                    ImportedPlanData[30:41,3],
                    ImportedPlanData[30:41,4],
                    ImportedPlanData[30:41,5],
                    ImportedPlanData[30:41,6],
                    ImportedPlanData[30:41,7])
YearPlanVector <- c(ImportedPlanData[86,2],
                    ImportedPlanData[86,3],
                    ImportedPlanData[86,4],
                    ImportedPlanData[86,5],
                    ImportedPlanData[86,6],
                    ImportedPlanData[86,7])
PlanVector_2014 <- c(ImportedPlanData[2:13,8])

# The data is saved as a vector and needs to be converted into a time series
TotalAsIs<- ts(TotalAsIsVector , start=c(2008,1), 
               end=c(2013,12), frequency=12)
WugeAsIs <- ts(WugeAsIsVector, start=c(2008,1), 
               end=c(2013,12), frequency=12)
YearAsIs <- ts(YearAsIsVector, start=c(2008,1), 
               end=c(2013,12), frequency=12)
TotalAsIs_2014 <- ts(TotalAsIsVector_2014, start=c(2014,1), 
                     end=c(2014,12), frequency=12)
TotalPlan <- ts(PlanVector , start=c(2008,1), 
                end=c(2013,12), frequency=12)
WugePlan <- ts(WugePlanVector, start=c(2008,1), 
               end=c(2013,12), frequency=12)
YearPlan <- ts(YearPlanVector, start=c(2008,1), 
               end=c(2013,12), frequency=12)
TotalPlan_2014 <- ts(PlanVector_2014, start=c(2014,1), 
                     end=c(2014,12), frequency=12)

#check out the data to make sure it's right
str(TotalAsIs)
str(WugeAsIs)
str(YearAsIs)
str(TotalAsIs_2014)
str(TotalPlan)
str(WugePlan) 
str(YearPlan)
str(TotalPlan_2014)

#EDA code
#code from WWBBFH analysis
# Test the correlation between As Is and Plan data in order to test how exact 
# the planning is. 
# Correlation is a measure of linear relationship between two variables. 

TotalCor <- cor(TotalAsIs, TotalPlan )
WugeCor <- cor(WugeAsIs, WugePlan)



# The results show a very high planning accuracy.

#fit a linear model to Total AsIs
WugeAsIs_lm <- lm(WugeAsIs ~ WugePlan , data = WugeAsIs)
summary(WugeAsIs_lm)

#fit the linear model TotalAsIs_lm to a time series
WugeAsIs_tslm <- tslm(WugeAsIs ~ TotalPlan)
summary(WugeAsIs_lm)

#use STL to breakdown seasonality, trend, and remainder
WugeAsIs_stl <- stl(WugeAsIs, s.window=5)


#review STL results on a plot

plot(WugeAsIs_stl, col="black", main="WugeAsIs_stl")


#look at trend for Wuge
par(mfrow=c(1,1), mar=c(6,4,1,1))
plot(WugeAsIs_stl$time.series[,"trend"], col="blue", main="Wuge Trend")

#compare the seasonality.  "It only makes sense to do 
#this if the seasonality componant as the trend looks almost identical and 
#the remainder is then randomly spread."

monthplot(WugeAsIs_stl$time.series[,"seasonal"], main="Monthly Seasonality", ylab="Seasonal")


## Correlation with external indicators

#The indicators are as follows:
  
#Monthly Change in Export Price Index (CEPI)
#Monthly Satisfaction Index (SI) government based data
#Average monthly temperatures in Chulwalar
#Monthly births in Chulwalar
#Monthly Satisfaction Index (SI) external index 
#Yearly Exports from Urbano
#Yearly number of Globalisation Party members in Chulwalar
#Monthly Average Export Price Index for Chulwalar
#Monthly Producer Price Index (PPI) for Etel in Chulwalar
#National Holidays
#Chulwalar Index (Total value of all companies in Chulwalar)
#Monthly Inflation rate in Chulwalar
#Proposed spending for National Holidays
#Influence of National Holiday

#The indicators will be converted into individual  vectors and subsequently converted into time series. The correlation of the indicators will then be tested against the As Is exports for Chulwalar. 

# Monthly Change in Export Price Index (CEPI)
CEPIVector <- c(ImportedIndicators[2:13,2],ImportedIndicators[2:13,3],ImportedIndicators[2:13,4],ImportedIndicators[2:13,5],ImportedIndicators[2:13,6],ImportedIndicators[2:13,7])
CEPI <- ts(CEPIVector , start=c(2008,1), end=c(2013,12), frequency=12)
plot(CEPI, main="CEPI")

WugeCEPICor <- cor(WugeAsIs, CEPI)

# Monthly Satisfaction Index (SI) government based data
SIGovVector <- c(ImportedIndicators[16:27,2],ImportedIndicators[16:27,3],ImportedIndicators[16:27,4],ImportedIndicators[16:27,5],ImportedIndicators[16:27,6],ImportedIndicators[16:27,7])
SIGov <- ts(SIGovVector , start=c(2008,1), end=c(2013,12), frequency=12)
plot(SIGov, main="SIGov")

WugeSICor <- cor(WugeAsIs, SIGov)

# Average monthly temperatures in Chulwalar
TemperatureVector <- c(ImportedIndicators[30:41,2],ImportedIndicators[30:41,3],ImportedIndicators[30:41,4],ImportedIndicators[30:41,5],ImportedIndicators[30:41,6],ImportedIndicators[30:41,7])
Temperature <- ts(TemperatureVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(Temperature, main="Temperature")

WugeTempCor <- cor(WugeAsIs, Temperature)

# Monthly births in Chulwalar 
BirthsVector <- c(ImportedIndicators[44:55,2],ImportedIndicators[44:55,3],ImportedIndicators[44:55,4],ImportedIndicators[44:55,5],ImportedIndicators[44:55,6],ImportedIndicators[44:55,7])
Births <- ts(BirthsVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(Births, main="Births")

WugeBirthCor <- cor(WugeAsIs, Births)

# Monthly Satisfaction Index (SI) external index 
SIExternVector <- c(ImportedIndicators[58:69,2],ImportedIndicators[58:69,3],ImportedIndicators[58:69,4],ImportedIndicators[58:69,5],ImportedIndicators[58:69,6],ImportedIndicators[58:69,7])
SIExtern <- ts(SIExternVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(SIExtern, main="SIExtern")

WugeSIECor <- cor(WugeAsIs, SIExtern)

# Yearly exports from Urbano
UrbanoExportsVector <- c(ImportedIndicators[72:83,2],ImportedIndicators[72:83,3],ImportedIndicators[72:83,4],ImportedIndicators[72:83,5],ImportedIndicators[72:83,6],ImportedIndicators[72:83,7])
UrbanoExports <- ts(UrbanoExportsVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(UrbanoExports, main="UrbanoExports")

WugeExpCor <- cor(WugeAsIs, UrbanoExports)

# Yearly number of Globalisation Party members in Chulwalar
GlobalisationPartyMembersVector <- c(ImportedIndicators[86:97,2],ImportedIndicators[86:97,3],ImportedIndicators[86:97,4],ImportedIndicators[86:97,5],ImportedIndicators[86:97,6],ImportedIndicators[86:97,7])
GlobalisationPartyMembers <- ts(GlobalisationPartyMembersVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(GlobalisationPartyMembers, main="GlobalisationPartyMembers")

WugeGloCor <- cor(WugeAsIs, GlobalisationPartyMembers)

# Monthly Average Export Price Index for Chulwalar
AEPIVector <- c(ImportedIndicators[100:111,2],ImportedIndicators[100:111,3],ImportedIndicators[100:111,4],ImportedIndicators[100:111,5],ImportedIndicators[100:111,6],ImportedIndicators[100:111,7])
AEPI <- ts(AEPIVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(AEPI, main="AEPI")

WugeAEPICor <- cor(WugeAsIs, AEPI)

# Monthly Producer Price Index (PPI) for Etel in Chulwalar
PPIEtelVector <- c(ImportedIndicators[114:125,2],ImportedIndicators[114:125,3],ImportedIndicators[114:125,4],ImportedIndicators[114:125,5],ImportedIndicators[114:125,6],ImportedIndicators[114:125,7])
PPIEtel <- ts(PPIEtelVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(PPIEtel, main="PPIEtel")

WugePPICor <- cor(WugeAsIs, PPIEtel)

# National Holidays
NationalHolidaysVector <- c(ImportedIndicators[170:181,2],ImportedIndicators[170:181,3],ImportedIndicators[170:181,4],ImportedIndicators[170:181,5],ImportedIndicators[170:181,6],ImportedIndicators[170:181,7])
NationalHolidays <- ts(NationalHolidaysVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(NationalHolidays, main="NationalHolidays")

WugeNatCor <- cor(WugeAsIs, NationalHolidays)

# Chulwalar Index (Total value of all companies in Chulwalar)
ChulwalarIndexVector <- c(ImportedIndicators[128:139,2],ImportedIndicators[128:139,3],ImportedIndicators[128:139,4],ImportedIndicators[128:139,5],ImportedIndicators[128:139,6],ImportedIndicators[128:139,7])
ChulwalarIndex <- ts(ChulwalarIndexVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(ChulwalarIndex, main="ChulwalarIndex")

WugeIdxCor <- cor(WugeAsIs, ChulwalarIndex)

# Monthly Inflation rate in Chulwalar 
InflationVector <- c(ImportedIndicators[142:153,2],ImportedIndicators[142:153,3],ImportedIndicators[142:153,4],ImportedIndicators[142:153,5],ImportedIndicators[142:153,6],ImportedIndicators[142:153,7])
Inflation <- ts(InflationVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(Inflation, main="Inflation")

WugeInfCor <- cor(WugeAsIs, Inflation)


# Proposed spending for Independence day presents
IndependenceDayPresentsVector <- c(ImportedIndicators[156:167,2],ImportedIndicators[156:167,3],ImportedIndicators[156:167,4],ImportedIndicators[156:167,5],ImportedIndicators[156:167,6],ImportedIndicators[156:167,7])
IndependenceDayPresents <- ts(IndependenceDayPresentsVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(IndependenceDayPresents, main="IndependenceDayPresents")

WugeIndCor <- cor(WugeAsIs, IndependenceDayPresents)

# Influence of National Holidays :
#This indicator is an experiment where the influence of National Holidays is 
#extended into the months leading up to the holiday. However later tests show 
#that this indicator is no better for forecasting than the orignial National 
#Holidays indicator.  

InfluenceNationalHolidaysVector <- c(ImportedIndicators[184:195,2],ImportedIndicators[184:195,3],ImportedIndicators[184:195,4],ImportedIndicators[184:195,5],ImportedIndicators[184:195,6],ImportedIndicators[184:195,7])
InfluenceNationalHolidays <- ts(InfluenceNationalHolidaysVector, start=c(2008,1), end=c(2013,12), frequency=12)
plot(InfluenceNationalHolidays, main="InfluenceNationalHolidays")

WugeIndCor2 <- cor(WugeAsIs, InfluenceNationalHolidays)
par(mfrow=c(1,1), mar = c(12,6,4,2) + 0.1, mgp = c(4,2,0))
Correlations <- c(WugeCEPICor, WugeExpCor, WugeSICor, WugeSIECor, WugeBirthCor, WugeTempCor, WugeGloCor, WugeAEPICor, WugePPICor, WugeNatCor, WugeIdxCor, WugeInfCor, WugeIndCor, WugeIndCor2 )
ParamNames <- c("Mo. Change in Export Price", "Mo. Avg. Export Price", "Mo.Satisfaction Idx - Govt", "Mo. Satisfaction Idx - Ext.", "Mo. Births", "Mo. Avg. Temperature", "Yearly Exports", "Yearly No. of GP Members", "Mo. Producer Price Idx", "National Holidays", "Chulwalar Idx", "Mo. Avg. Inflation Rate", "Proposed Ind. Day Gifts", "Influence of Ind. Day Gifts")
plot(Correlations, main="Correlations for Wuge Plant", xaxt="n", pch=19,col="blue", ylim=c(-1, 1), xlab = "", ylab="Correlation")
axis(1, labels = ParamNames, at=1:14, las=2)

# Check that the data import has worked
str(CEPIVector)
str(SIGovVector)  
str(TemperatureVector) 
str(BirthsVector)
str(SIExternVector)
str(UrbanoExportsVector) 
str(GlobalisationPartyMembersVector)
str(AEPIVector) 
str(PPIEtelVector) 
str(NationalHolidaysVector) 
str(ChulwalarIndexVector)
str(InflationVector) 
str(IndependenceDayPresentsVector)

par(mfrow=c(1,1))
#Forecasting work
# Forecasting models with smoothing and related approaches
#Exponential Smoothing uses past values to calculate a forecast. 
#The strength with which each value influences the forecast is 
#weakened with help of a smoothing parameter. Thus we are dealing 
#with a weighted average, whose values fade out the longer ago 
#they were in the past. 

#The Akaike's Information Criterion(AIC/AICc) or the 
#Bayesian Information Criterion (BIC) should be at minimum.


## Simple expontential smoothing 
par(mfrow=c(1,1))
Model_ses <- ses(WugeAsIs, h=12)
summary(Model_ses)
plot(Model_ses, plot.conf=FALSE, ylab="Wuge Exports Chulwalar", xlab="Year", main="", fcol="white", type="o")
lines(fitted(Model_ses), col="green", type="o")
lines(Model_ses$mean, col="blue", type="o")
legend("topleft",lty=1, col=c(1,"green"), c("data", expression(alpha == 0.497)),pch=1)

## Holt's linear trend method   
# Holt added to the model in order to forecast using trends as well. 
# For this it is necessary to add a beta, which determines the trend. 
# If neither alpha nor beta is stated, both parameters will be optimised 
# using ets(). The trend is exponential if the intercepts(level) and the 
# gradient (slope) are multiplied with eachother. The values are worse. As
# the Beta was very low in the optimisation, the forecast is very similar 
# to the ses() model. 

Model_holt_1 <- holt(WugeAsIs,h=12)
summary(Model_holt_1)
plot(Model_holt_1)


# expoential trend

Model_holt_2<- holt(WugeAsIs, exponential=TRUE,h=12)
summary(Model_holt_2)
plot(Model_holt_2)


## Dampened trends
# As such simple trends tend to forecast the future too positively, we have 
# added a dampener. This also works for exponential trends. We also plot 
# the level and slope individually for each model.

Model_holt_3 <- holt(WugeAsIs, damped=TRUE,h=12)
summary(Model_holt_3)
plot(Model_holt_3)

Model_holt_4 <- holt(WugeAsIs, exponential=TRUE, damped=TRUE,h=12)
summary(Model_holt_4)
plot(Model_holt_4)

# level and slope can be plotted individually for each model. 
plot(Model_holt_1$model$state)
plot(Model_holt_2$model$state)
plot(Model_holt_3$model$state)
plot(Model_holt_4$model$state)

plot(Model_holt_1, plot.conf=FALSE, ylab=" Wuge Exports Chulwalar  )", xlab="Year", main="", fcol="white", type="o")
lines(fitted(Model_ses), col="purple", type="o")
lines(fitted(Model_holt_1), col="blue", type="o")
lines(fitted(Model_holt_2), col="red", type="o")
lines(fitted(Model_holt_3), col="green", type="o")
lines(fitted(Model_holt_4), col="orange", type="o")
lines(Model_ses$mean, col="purple", type="o")
lines(Model_holt_1$mean, col="blue", type="o")
lines(Model_holt_2$mean, col="red", type="o")
lines(Model_holt_3$mean, col="green", type="o")
lines(Model_holt_4$mean, col="orange", type="o")
legend("topleft",lty=1, col=c(1,"purple","blue","red","green","orange"), c("data", "SES","Holts auto", "Exponential", "Additive Damped", "Multiplicative Damped"),pch=1)


## Holt-Winter's seasonal method   
# Holt and Winters have expanded Holt's model further to include
# the seasonality aspect. The parameter gamma, which is for smoothing 
# the seasonality, was added to achieve this. The values are better than 
# the models without seasonality. This is logical, since the data is strongly
# influenced by seasonality.  In the following model, none of the parameters 
# are given so that they will be optimised automatically. There are two models: 
# one using an additive error model method and one using a multiplicative error model. 
# The additive model gives slightly better results than the multiplicative model.

Model_hw_1 <- hw(WugeAsIs ,seasonal="additive",h=12)
summary(Model_hw_1)
plot(Model_hw_1)

Model_hw_2 <- hw(WugeAsIs ,seasonal="multiplicative",h=12)
summary(Model_hw_2)
plot(Model_hw_2)

plot(Model_hw_1, ylab="Wuge Exports", plot.conf=FALSE, type="o", fcol="white", xlab="Year")
lines(fitted(Model_hw_1), col="red", lty=2)
lines(fitted(Model_hw_2), col="green", lty=2)
lines(Model_hw_1$mean, type="o", col="red")
lines(Model_hw_2$mean, type="o", col="green")
legend("topleft",lty=1, pch=1, col=1:3, c("data","Holt Winters' Additive","Holt Winters' Multiplicative"))

#Damp the Holt-Winter models
Model_hw_3 <- hw(WugeAsIs ,seasonal="additive", damped = TRUE,h=12)
Model_hw_4 <- hw(WugeAsIs ,seasonal="multiplicative", damped = TRUE,h=12)

#create plot to show all HW lines

HW1 <- hw(WugeAsIs, seasonal = "additive")
HW2 <- hw(WugeAsIs, seasonal = "multiplicative")
HW3 <- hw(WugeAsIs, seasonal = "additive", damped = TRUE)
HW4 <- hw(WugeAsIs, seasonal = "multiplicative", damped = TRUE)



# In order to use the results later, they need to be converted into point forecasts.
Model_hw_1_df <-as.data.frame(Model_hw_1) 
Model_hw_1_PointForecast <- ts(Model_hw_1_df$"Point Forecast", start=c(2014,1), end=c(2014,12), frequency=12)
Model_hw_1_PointForecast
Model_hw_2_df <-as.data.frame(Model_hw_2) 
Model_hw_2_PointForecast <- ts(Model_hw_2_df$"Point Forecast", start=c(2014,1), end=c(2014,12), frequency=12)
Model_hw_2_PointForecast

#measures of quality
QualityNames <- c("AIC", "AICc", "BIC")
SESMeasures <- c(1981.370, 1981.544, 1985.923)
HoltAdditive <- c(1984.278, 1984.875, 1993.385)
HoltExponential <- c(1992.221, 1992.818, 2001.327)
HoltAddDamp <- c(1986.510, 1987.419, 1997.893)
HoltExpDamp <- c(1995.499, 1996.408, 2006.883)
HWAdditive <- c(1910.881, 1920.771, 1947.307)
HWExponential <- c(1910.516, 1920.407, 1946.942)
HWAdditiveDamp <- c(1919.078, 1930.411, 1957.781)
HWExponentialDamp <- c(1919.692, 1931.025, 1958.395)

#info for error information / presentation
ErrorNames <- c("ME", "RMSE", "MAE", "MPE", "MAPE", "MASE", "ACF1")
Methods <- c("SES", "Holt Add.", "Holt Mult.", "Holt Add. w/ Damp", "Holt Mult. w/ Damp", "HW's Add.", "HW's Mult.", "HW's Add. w/ Damp", "HW's Mult. w/ Damp")
ME <- c(14472.94, -7995.956, 2854.209, 3587.916, 6750.188, -63.33698, -4709.478, 7939.675, 8344.003)
RMSE <- c(108380.5, 107561.6, 107620.5, 107735, 108003.1, 54690.67, 54484.25, 57095.66,58602.93)
MAE <- c(79591.45, 83814.52, 83306.45, 82222.7, 81857.82, 43861.37, 44908.38, 46556.41,47747.06)
MPE <- c(-0.2985412, -4.528997, -2.476202, -2.590547, -1.869407, -1.398017, -2.31886, 0.4534177,0.2032803 )
MAPE <- c(13.80368, 15.17678, 14.72429, 14.64643,14.43078, 7.97784, 8.351825, 8.459067,8.478591)
MASE <- c(0.9695715, 1.021016, 1.014827, 1.01625, 0.9971801, 0.5343129, 0.5470674, 0.5671435,0.5816478)
ACF1 <- c(0.02262988, 0.05692744, 0.1136907, 0.06340987, 0.065998, -0.1775947, -0.309649, -0.1828394,-0.27984804)

#info for Holt Winters' errors
Methods2<- c("HW's Add.", "HW's Mult.", "HW's Add. w/ Damp", "HW's Mult. w/ Damp")
HWMAE<- c(43861.37, 44908.38, 46556.41,47747.06)
HWMAPE <- c(7.97784, 8.351825, 8.459067,8.478591)
HWErrors <- data.frame(Methods2, HWMAE, HWMAPE)
names(HWErrors) <- c("Method", "MAE", "MAPE")


#Residuals
#create new set of holt winters models levels, slopes, and seasonals

states <- cbind(Model_hw_1$model$states[,1:3], Model_hw_2$model$states[,1:3])
colnames(states) <- c("Level" , "Slope", "Seasonal", "Level" , "Slope", "Seasonal")

```

